Multiscale Modeling of the Paradoxical Roles of the Immune System in EMT-Mediated Cancer

### 2019-02-05 Paper/Paper.tex
add tcga figure

\documentclass{article}
\usepackage{outline}
\usepackage{float}
\usepackage{graphicx}
\usepackage{color}
\usepackage{amsmath}
\usepackage{cite}
\graphicspath{ {Figs/} }

\begin{document}

 \title{Multiscale Modeling of the Paradoxical Roles of the Immune System in EMT-Mediated Cancer}
 \maketitle

\section*{Abstract}
%% Basic Introduction (1-2 sentences)
During tumorigenesis and throughout the lifetime of a tumor, the immune system is constantly engaged in complex interactions in and around the tumor.
%% More Detailed Background (2-3 sentences)
These interactions are influenced by the tumor microenvironment (TME) and in particular by the cellular process of epithelial-to-mesenchymal transition (EMT) in an as-yet unexplained manner.
In turn, both the tumor and the immune system exert substantial influence over the TME in competition to influence the fate of the host.
%% General Problem (1 sentence)
Determining the critical factors and mechanisms involved remains a crucial task for cancer immunotherapy.
%% Summarizing Main Result (1 sentence)
Here, we seek to understand how interplay between the tumor, the immune system, and EMT can determine the outcome of cancer.
%% The Main Result (2-3 sentences)
We developed a multiscale agent-based model that describes tumor-immune interactions and how EMT modulates these.
Exploration of the model revealed that increasing the growth arrest of mesenchymal cells in some circumstances results in a more tumor-friendly environment and sometimes less.
In contrast, enhancing the immune evasiveness of mesenchymal cells always leads to a more tumor-friendly microenvironment.
Moreover, encouraging the TME towards a pro-EMT state also benefits the tumor.
%% Putting into a More General Context (1-2 sentences)
Together, these results indicate that the joint regulation of the TME by the immune system and the tumor itself provides possible targets for future cancer therapies, in particular cancers such as pancreatic cancer in which EMT is speculated to play a role.

@@

\section{Introduction}
Immunotherapy is slowly revolutionizing how we treat cancer.
Year-by-year, new breakthroughs in the field produce new, effective drugs and therapies that have a significant impact on patient health and survival\cite{pardoll2012blockade}\cite{restifo2012adoptive}.
% A More examples? Reviews or non-reviews? Explain some of the examples?
%AM I will add a few 
There are myriad untapped avenues by which scientists can further explore the role of the immune system in regulating and combatting cancer; in addition to direct cell-cell interactions between immune and cancer cells, the means by which immune cells interact with and modulate the tumor microenvironment (TME) is an important research area. Immune cells can be understood as one constituent of the TME: responding to the environment's inflammatory signals, fighting the mutated cells present, and then restoring homeostasis in that local context\cite{de2006paradoxical}.
Understanding each step in this immune cascade in the context of the TME will be necessary in order to fully explore and exploit all options of immunotherapy.

@@

\subsection{Heat Map of MIE vs MGA}
To further draw out the contrast between these two mesenchymal properties, we compare Time to Cancer for many different pairs of values for MIE and MGA.
The heat map in Figure \ref{fig:MIEvsMGA} concisely encodes the relationships we have been observing in the previous sections.
Regardless of the value of MGA, increasing MIE leads to decreased Time to Cancer.
On the other hand, for any value of MIE, then there is an MGA value that results in the longest Time to Cancer.
For the model, this optimal value increases with MIE.

@@

>>>>

\documentclass[11pt, a4paper, preprint]{article}
\usepackage{outline}
\usepackage{float}
\usepackage{graphicx}
\usepackage{color}
\usepackage{amsmath}
\usepackage{cite}
\graphicspath{ {Figs/} }

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{authblk}
\newcommand{\tcr} { \textcolor{red} }

\title{Multiscale Modeling of the Paradoxical Roles of the Immune System in EMT-Mediated Cancer}
\author{}

\renewcommand\Authands{ and }

\begin{document}
\maketitle

\section*{Abstract}
%% Basic Introduction (1-2 sentences)
Preceding tumorigenesis and throughout the lifetime of a tumor, the immune system is engaged in complex sets of interactions in and around the tumor.
%% More Detailed Background (2-3 sentences)
These are regulated by the tumor microenvironment (TME) in myriad ways. Moreover,  epithelial cell plasticity impacts disease incidence and progression via processes of epithelial-to-mesenchymal transition (EMT).
In turn, both the tumor and the immune system exert substantial influence over the TME in competition to influence the fate of the host.
%% General Problem (1 sentence)
Determining the factors and mechanisms underpinning these complex interactions remains an important task for cancer immunotherapy.
%% Summarizing Main Result (1 sentence)
%Here, we seek to understand how interplay between the tumor, the immune system, and EMT can determine the outcome of cancer.
%% The Main Result (2-3 sentences)
We developed a multiscale agent-based model to describe the interplay between three components: the tumor (or pre-tumor cells), the immune system, and EMT.
The model revealed a non-monotonic relationship between the cell cycle times associated with the mesencymal phenotype and the probability of tumorigenesis, identifying a critical growth rate that maximizes cancer-free survival. In comparison, increasing the immune-evasiveness of the mesenchymal phenotypes always led to poorer prognosis. Moreover, encouraging
%\tcr
the TME
towards a pro-EMT state also benefits the tumor.
%% Putting into a More General Context (1-2 sentences)
Together, these results indicate that the joint regulation of the TME by the immune system and the tumor itself provides possible targets for future cancer therapies, in particular cancers such as pancreatic cancer in which EMT is speculated to play a role.

@@

\section{Introduction}
Immunotherapy is slowly revolutionizing cancer therapy. New breakthroughs in the field begin to produce new therapies that have had a significant impact on patient health and survival\cite{pardoll2012blockade}\cite{restifo2012adoptive}.
% A More examples? Reviews or non-reviews? Explain some of the examples?
%AM I will add a few 
%There are myriad untapped avenues by which scientists can further explore the role of the immune system in regulating and combatting cancer;
In addition to direct cell-cell interactions between immune and cancer cells, the means by which immune cells interact with and modulate the tumor microenvironment (TME) is an important research area. Immune cells can be understood as one constituent of the TME: responding to the environment's inflammatory signals, fighting the mutated cells present, and then restoring homeostasis in that local context\cite{de2006paradoxical}.
Understanding each step in this immune cascade in the context of the TME will be necessary in order to fully explore and exploit all options of immunotherapy.

@@

% \subsection{Heat Map of MIE vs MGA}
\subsection{Analysis of TCGA data in comparison with model predictions suggests that mesenchymal phenotypes reduce cancer-free survival probability}

To further analyze the effects on tumorigenesis of the mesenchymal properties of immune evasion (MIE) and growth arrest (MGA), we study the joint density plot detailing the effects of each of these parameters on the Time to Cancer (Fig. \ref{fig:MIEvsMGA}). We found that over the full range of values of MGA considered, increasing the MIE decreases the Time to Cancer. However, for any given value of mesenchymal immune evasion, there is a value of mesenchymal growth arrest that maximizes the Time to Cancer. Moreover, this optimal value increases with MIE.

@@

To compare these model predictions with experimental studies, we performed analysis of data from The Cancer Genome Atlas (TCGA) database to study the effects of immune and EMT interactions on prognosis of cancers for which inflammation is known to play an important role, such as colonic or pancreatic cancers (REFS). We selected pancreatic cancer to investigate, and used gene ontologies as a measure of the XXXX of effects of inflammatory or EMT processes on survival. In Fig. \ref{fig:tcga} we plot Kaplan-Meier curves for

\begin{figure}[H]
\center
\frame{\includegraphics[width=\columnwidth]{FigTCGA.pdf}}
\caption{......r.}
\label{fig:tcga}
\end{figure}
